### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_1.Rnw
package skeleton for yuimadoc package, not yet working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@11 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Introduction}

\def\yuima{{\bf yuima}}

The package $\yuima$ provides 
an object-oriented programming environment 
for simulation and statistical inference 
for stochastic processes by R. 
On this framework, 
$\yuima$ also supplies various functions 
to execute simulation and statistical analysis. 
Both categories of procedures may depend each other. 
Statistical inference often requires a simulation technique 
as a subroutine, and a certain simulation method 
needs to fix a tuning parameter by applying 
a statistical methodology. 
It is especially the case of stochastic processes 
because most of expected values involved 
do not admit an explicit expression. 
%It is therefore natural to 
$\yuima$ facilitates comprehensive, systematic approaches 
to the solution.

Stochastic differential equations are 
commonly used 
to model random evolution along continuous or 
practically continuous time, such as 
the random movements of a stock price. 
Theory of statistical inference for 
stochastic differential equations already has a fairly long history, 
more than three decades, but it is still developing quickly new 
methodologies and expanding the area. 
The formulas produced by the theory are usually very sophisticated, 
which makes it difficult for standard users not necessarily 
familiar with this field to enjoy utilities. 
For example, the asymptotic expansion method for computing 
option prices (i.e., expectation of an irregular functional of 
a stochastic process) provides precise approximation values 
instantaneously, taking advantage of the analytic approach, 
but the formula has a long expression like more than one page! 
$\yuima$ delivers up-to-date methods as a package 
onto the desk of the user working 
with simulation and/or statistics for stochastic differential equations.

Sampled data from a continuous-time process features 
the time stamps as well as the positions of the object. 
It is requiring a new theory of estimation. 
$\yuima$ framework can apply multi-dimensional time stamps 
of tick data and provides diverse functions handling such kind data 
to support statistical analysis.