### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_2.Rnw

\section{Yuima object}
The main object in our framework is the \code{yuima} object.
The \code{yuima} object contains slots named \code{model}, \code{data},
\code{sampling}, and \code{characteristic}.

\begin{center}
\begin{tabular}{|c|}
\hline
\colorbox{black}{\textcolor{white}{  \code{yuima} \bf object}}\\
\hline
\hline
\code{model}\\
\hline
\code{data}\\
\hline
\code{sampling}\\
\hline
\code{characteristic}\\
\hline
\end{tabular}
\end{center}

The slot \code{data} will contain
a time series data and the slot \code{sampling} contains a description of
sampling methods and/or parameters of the data collection(e.g. poisson sampling with rate $p=0.1$.).

The \code{yuima.model} object is intended to be a description of model 
independent of the actual data. 
For example, in model selection or hypothesis testing for SDEs, 
it is possible to have different models to fit to the same set of data.

On the contrary, for the same model specified in \code{yuima.model}, 
different time series can be fitted. 
In this case the vector \code{model} has length 1, and 
the vector of \code{data} has length greater than 1.

The \code{yuima.characteristic} object is prepared for specifying the characteristics of the \code{yuima} object such as
the number of equations in the model and data, time and space unit scale for simulation or data.
For now, this slot is mainly for keeping consistency of the other three slots.
%For example, the number of pathes specified in the \code{yuima.model} object and the actual data in \code{yuima.data} is 
%not the same,

\code{setYuima} is a constructor for \code{yuima} class, 
which does not initialize but only creates four NULL slots:

\begin{itemize}
\item{data:}{ an object of class \code{yuima.data}}
\item{model:}{ an object of class \code{yuima.model}}
\item{sampling:}{ an object of class \code{yuima.sampling}}
\item{characteristic:}{ an object of class \code{yuima.characteristic}}
\end{itemize}
%We show how to construct \code{yuima} object in the next section.

%%Note that \code{yuima.sampling} is not implemented yet.

%-----------------------------------
 \begin{figure}[!htb]
  \centering
  \includegraphics[width=10cm,clip]{./EPS/functionDefinition.eps}
  \caption{function relation}
 \end{figure}
 %-----------------------------------
%-----------------------------------
 \begin{figure}[!htb]
  \centering
  \includegraphics[width=5cm,clip]{./EPS/classDefinition.eps}
  \caption{class definition}
 \end{figure}
 %-----------------------------------

\subsection{yuima.model}
\code{yuima.model} specifies the SDE of interest. The constructor \code{setModel} is
used to give a mathematical description of the stochastic differential equation. 
All functions in the package are assumed to get as much information
as possible from the model instead of replicating the same code everywhere.
If there are missing pieces of information, we may change or extend the description
of the model.
\\
\\
An object of \code{yuima.model} contains several slots listed below. 
To see inside its structure, we use the \R{} command \code{str}.
\begin{itemize}
\item \code{drift} is an \R{} expression which contains the drift specification.
\item \code{diffusion} is itself a list of 1 slot which describes the diffusion
  coefficient relative to first noise.

\item \code{parameter} which is a short name for ``parameters'' which 
  is a list of objects.

\item \code{all} contains the names of all the parameters found 
   in the diffusion and drift coefficient.
\item \code{common} contains the names of the parameters in common between the drift and diffusion coefficients.
\item \code{diffusion} contains the parameters belonging to the diffusion coefficient.
\item \code{drift} contains the parameters belonging to the drift coefficient.
\item \code{solve.variable} contains a vector of variable names, each element corresponds to the
   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.

\item \code{state.variable} and \code{time.variable}, by default,
  are assumed to be $x$ and $t$ but the user can freely choose them.
  The \code{yuima.model} function assumes that the user either use default
  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
  the rest of the symbols are considered parameters and distributed accordingly
  in the \code{parameter} slot.

\item \code{noise.number} indicates the number of sources of noise.

\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
  stochastic differential equations.

\item \code{dimension} reports the dimensions of the parameter space. It is a list of the
  same length of \code{parameter} with the same names.

\end{itemize}
We show an example to describe the following simple stochastic differential equation

\begin{equation}
\mbox{d}X_t=3X_t\mbox{d}t+\mbox{d}W_t.
\end{equation}

<<print=FALSE,echo=TRUE>>=
library(yuima)
@

<<echo=TRUE>>=
mod <- setModel(drift="3*x")
str(mod)
@

\subsection{yuima.data}
\code{yuima.data} is a class to represent the data to be used. The constructor \code{setData} is
used to give an observed data to the \code{yuima.data} object. \code{get.zoo.data} returns the \code{zoo.data} field of 
an \code{yuima.data} object.
As usual class methods, \code{yuima.data} provides \code{length}, \code{plot} and 
\code{dim} functions to get necessary information.

 The \code{zoo.data} slot may contain different kind of observed data. 
 In the \R{} system, there are different kinds of objects to describe time series data.
 Here is a very short list of possibilities
 \begin{center}
 \begin{tabular}{r l}
 \code{ts}, \code{mts}: & regularly spaced time series 1-dim or multidim\\
 \code{its}: & irregularly spaced time series\\
 \code{tseries}: & time index is real time of financial markets \\
 \code{zoo}: & more flexibility  in the management of the index of the time series.\\ 
 \end{tabular}
 \end{center}
 Because \R{} is used in many contexts, including finance, and users adopt one of 
 the above times series objects because they are used to it, 
 our framework accepts any of the above as observed data and treat the inference accordingly.
 \\
 \\
 An object of \code{yuima.data} contains several slots listed below. 
 To see inside its structure we use the \R{} command \code{str}.
 \begin{itemize}
 \item \code{original.data} is a copy of the original data. (it can be any type of data.)
 \item \code{zoo.data} is an bundled \code{zoo} version object of class \code{yuima.data} or \code{yuima}.
 \end{itemize}

 We show an example of \code{yuima.data} object.

<<print=FALSE,echo=TRUE>>=
library(zoo)
@

<<fig=TRUE,echo=TRUE,width=7,height=7,results=hide>>=
X <- ts(matrix(rnorm(200),100,2))
mydata <- setData(X)
str(get.zoo.data(mydata))
#plot(get.zoo.data(mydata))
dim(mydata)
length(mydata)
plot(mydata)
@

Then, we show how to construct \code{yuima} object with \code{yuima.data} and \code{yuima.model} object.

We first construct a two-dimensional homogeneous diffusion process:
\begin{eqnarray*}
\mbox{d} X_{1t} &=&
- 3 X_{1t} \mbox{d}t + \mbox{d} W_{1t}, \\
\mbox{d}X_{2t} &=&
-(X_{1t} + 2 X_{2t}) \mbox{d}t + 0.5 \mbox{d} W_{1t} + \mbox{d}W_{2t}.
\end{eqnarray*}
<<echo=TRUE>>=
# we set the drift coefficient
a <- c("-3*x1","-x1-2*x2")
# and also the diffusion coefficient
b <- matrix(c("1","0.5","0","1"),2,2)
# Then set
mod2 <- setModel(drift = a, diffusion = b, solve.variable = c("x1","x2"))
str(mod2)
@

Then, we can combine the \code{yuima.model} and \code{yuima.data}
objects to a \code{yuima} object.
<<echo=TRUE>>=
yuima.obj <- setYuima(data=mydata,model=mod2)
str(yuima.obj)
@

\subsection{Realworld data}
In this subsection, we show a simple example of handling the realworld data in
the yuima framework.
<<fig=TRUE,echo=TRUE,width=7,height=7>>=
## Use TTR package to get real market data from Yahoo! Finance
require(TTR)
sbux<-getYahooData("SBUX", start=20080401, end=20090331)
cbou<-getYahooData("CBOU", start=20080401, end=20090331)
sd <- setData(list(sbux,cbou))
real.yuima <- setYuima(data=sd)
summary(real.yuima)
head(get.zoo.data(real.yuima)[[1]])
head(get.zoo.data(real.yuima)[[2]])

plot(get.zoo.data(real.yuima)[[1]]$High,main="SBUX_High")
@