### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_3.Rnw

\section{Simulation}
\subsection{Diffusion Process}
\code{simulate} is a function to solve SDE using the Euler-Maruyama method.
Here we illustrate how to generate a simulated path of the solution of
a given SDE. Let us consider a 3-dim SDE

 \begin{equation} \label{sabr}
 \begin{cases}
 & \mbox{d} X_t^1 = X_t^2 \left| X_t^1 \right|^{2/3} \mbox{d}W_t^1, \\
 & \mbox{d}X_t^2 = g(t)\mbox{d}X_t^3, \\
 & \mbox{d}X_t^3 = X_t^3( \mu  \mbox{d}t + \sigma (\rho \mbox{d}W_t^1 + \sqrt{1-\rho^2} 
   \mbox{d}W_t^2))
 \end{cases},
 \ \ \ \ (X_0^1, X_0^2, X_0^3) = (1.0,0.1,1.0)
 \end{equation}
 with $\mu = 0.1, \sigma = 0.2, \rho = -0.7 $ and $g(t) = 0.4 + (0.1 + 0.2t) e^{-2t}$,  for
 example, where $W = (W^1, W^2)$ is a 2-dim standard Brownian motion.
 The solution  $X = (X^1, X^2, X^3)$ takes values on $\mathbb{R}_+^3$.
 This is a stochastic integral equation defined as
 \begin{equation}
 X_t = X_0 +  \int_0^t b(s,X_s) \mbox{d}s + \int_0^t c(s,X_s) \mbox{d}W_s
 \end{equation}
 with 
 \[
  b(s,x) = \left( \begin{array}{@{\,}c@{\,}} 0 \\ g(s) \mu x^3 \\ \mu x^3 \end{array} \right),
 \ \ 
 c(s,x) = \left( \begin{array}{@{\,}cc@{\,}} x^2 | x^1|^{2/3} &  0 \\ 
  g(s)\sigma \rho x^3 &  g(s)\sigma \sqrt{1-\rho^2} x^3 \\ 
 \sigma \rho  x^3 & \sigma \sqrt{1-\rho^2} x^3 
 \end{array} \right)
 \]
 for $ x = (x^1, x^2, x^3)$. 
 The Euler-Maruyama method with discretization scheme
 \begin{equation}
 0 = \tau_0 < \tau_1 < \cdots < \tau_j < \tau_{j+1} < \cdots
 \end{equation}
 refers to an approximative solution given by
 \begin{equation}
 \tilde{X}_{\tau_{j+1}} = \tilde{X}_{\tau_j} + 
 b(\tau_j,\tilde{X}_{\tau_j})(\tau_{j+1} - \tau_j) + 
 c(\tau_j, \tilde{X}_{\tau_j} )(W_{\tau_{j+1}}- W_{\tau_j}), \ \ j \geq 0.
 \end{equation}
 If interval $\tau_{j+1} - \tau_j$ is independent to  $\{W_t\}_{t \geq \tau_j}$ for each $j$,
 $W_{\tau_{j+1}} - W_{\tau_j}$ has the same distribution as $\sqrt{\tau_{j+1} - \tau_j}N_j$,
 where $N_j$ is an iid sequence of standard normal variables.
 The usual discretization scheme is an equidistant one such as 
 \begin{equation}
 \tau_{j+1} - \tau_j = \text{Terminal}/\text{division}
 \end{equation}
 for all $j \geq 0$.

<<fig=TRUE,echo=TRUE,width=7,height=7>>=
 set.seed(123)

 mu <- 0.1
 sig <- 0.2 
 rho <- -0.7

 g <- function(t){
    return(0.4 + (0.1 + 0.2*t)* exp(-2*t))
 }

 f1 <- function(t, x1, x2, x3) {
    ret <- 0
    if(x1 > 0 && x2 > 0) ret <- x2*exp(log(x1)*2/3)
    return(ret)
 }

 f2 <- function(t, x1, x2, x3) {
     ret <- 0
     if(x3 > 0) ret <- rho*sig*x3
     return(ret)
 }

 f3 <- function(t, x1, x2, x3) {
     ret <- 0
     if(x3 > 0) ret <- sqrt(1-rho^2)*sig*x3
     return(ret)
 }

 diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
     "f2(t,x1,x2,x3)", "0", "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)

 sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
                       state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))

 # Change the parameters below if needed
 yuima.samp <- setSampling(Terminal=1,division=1000)

 # make yuima object
 yuima <- setYuima(model=sabr.mod, sampling=yuima.samp)

 # Solve SDEs using Euler-Maruyama method
 yuima <- simulate(yuima, xinit=c(1.0,0.1,1.0))
 plot(yuima)
@

 \subsection{Space-discretizing scheme}

 \code{simulate} provides also space-discretized Euler-Maruyama method to solve SDE.
 This is at the moment designed for 1 factor SDE, i.e., the case the driving
 Brownian motion $W$ is 1-dimensional. Here discretization scheme $\{\tau_j\}$ 
 is defined as

\begin{equation}
\tau_0 = 0, \ \ \tau_{j+1} = \inf\{ t > \tau_j; |W_t - W_{\tau_j}| = \epsilon \}
\end{equation}
for each $j \geq 0$. \code{simulate} takes
\begin{equation}
\epsilon^2 = \text{Terminal} / \text{division},
\end{equation}
which coinsides with the mean of the interval  $\tau_{j+1} - \tau_j$. 
This space-discretizing scheme is 3 times efficient than the usual time-discretizing
one in the sense of mean-squared error.

Let us consider the previous model (\ref{sabr}) with $\rho = -1$.

<<fig=TRUE,echo=TRUE,width=7,height=7>>=

set.seed(123)

rho <- -1

diff.coef.matrix <- c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
    "f2(t,x1,x2,x3)")

sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
                      state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))

yuima <- setYuima(model=sabr.mod, sampling=yuima.samp)

#yuima <- simulate(yuima)
yuima.sd <- simulate(yuima, space.discretized = TRUE, xinit=c(1.0,0.1,1.0))

#plot(yuima)
plot(yuima.sd)
@