### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_4.Rnw

\section{cce}
  cce is a function to estimate the covariance between two It\^o 
  processes when they are observed at discrete times nonsynchronously. 
  It can also apply to irregularly sampled one-dimensional data as a special case.

\subsection{Nonsynchronous covariance estimator}
Suppose that two It\^o processes are observed
only at discrete times in a nonsynchronous manner.
We are interested in estimating the covariance of the two processes accurately
in such a situation.
This type of problem arises typically in high-frequency financial time series.

Let $T \in (0,\infty)$ be a terminal time for possible observations. 
We consider a two dimensional It\^o process $(X^1,X^2)$  
satisfying the stochastic differential equations
\begin{eqnarray*}
\mbox{d}X_t^l &=&\mu_t^l \mbox{d}t + \sigma_t^l \mbox{d}W_t^l ,\quad
t\in[0,T]\\
X_0^l &=& x_0^l        \notag
\end{eqnarray*}
for $l=1,2$. 
Here $W^l$ denote standard Wiener processes with a 
progressively measurable correlation process 
$\mbox{d}\langle W_1 , W_2 \rangle_t = \rho_t \mbox{dt}$,  
$\mu^l_t$ and $\sigma_t^l$ are progressively measurable processes, 
and $x^l_0$ are initial random variables independent of $(W^1,W^2)$.  
Diffusion type processes are in the scope but this model 
can express more sophisticated stochastic structures.

The process $X^l$ is supposed to be observed at 
over the increasing sequence of times 
$T^{l,i}$ $(i\in\mathbb{Z}_{\geq0})$ starting at $0$, up to time T. 
Thus, the observables are $(T^{l,i},X^{l,i})$ with $T^{l,i}\leq T$. 
Each $T^{l,j}$ may be a stopping time, so possibly depends on 
the history of $(X^1,X^2)$ as well as the precedent stopping times.  
Two sequences of stopping times 
$T^{1,i}$ and $T^{2,j}$ are {\it nonsynchronous},  and 
irregularly spaced, in general. 
In particular, cce can apply to estimation of the quadratic variation of a single 
stochastic process sampled regularly/irregularly.

The parameter of interest is the quadratic covariation between $X^1$ and $X^2$:

\begin{equation}
\theta=
\langle X^1 , X^2 \rangle_T = \int_0^T \sigma_t^1 \sigma_t^2 \rho_t \mbox{d}t.
\end{equation}
The target variable $\theta$ is random in general.

It can be estimated with the nonsynchronous covariance estimator 
(Hayashi-Yoshida estimator) 
\begin{equation}
U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)
1_{\{ (T^{1,i-1},T^{1,i}] 
\bigcap (T^{2,j-1},T^{2,j}]  \neq \emptyset  \}}.
\end{equation}
That is, the product of any pair of increments $(X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)$ and
$(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)$ will make a contribution 
to the sum only when 
the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping 
with each other. 
%The estimator $U_n$ was proposed by Hayashi and Yoshida  and 
It is known that $U_n$ is consist and has asymptotically mixed normal 
distribution 
as $n\to\infty$ if the maximum length between two consecutive observing times 
tends to $0$. 
See \cite{hay-yos05,hay-yos04,hay-yos06,hay-yos08} for details.

\subsection{Example: data generation and estimation by yuima package}
We will demonstrate how to apply cce function to 
nonsynchronous high-frequency data by simulation. 
As an example, consider a two dimensional stochastic process 
$(X^1_t,X^2_t)$ satisfying the stochastic differential equation
\begin{equation}
\begin{split}
\mbox{d}X^1_t = \sigma_{1,t} \mbox{d}B^1_t, \\
\mbox{d}X^2_t = \sigma_{2,t} \mbox{d}B^2_t. 
\end{split}
\end{equation}
Here $B^1_t$ and $B^2_t$ denote two standard Wiener processes, 
however they are correlated as 
\begin{eqnarray}
B^1_t &=& W^1_t,\\
B^2_t &=&  \int_0^t \rho_s \mbox{d} W^1_s +
  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W^2_s,
\end{eqnarray}
where  $W^1_t$ and $W^2_t$ are independent Wiener processes,  and 
$\rho_t$ is the correlation function between $B^1_t$ and $B^2_t$. 
We consider $\sigma_{l,t},l=1,2$ and $\rho_t$ 
of the following form in this example:

\begin{eqnarray*}
 \sigma_{1,t} &=& \sqrt{1+t}, \\
 \sigma_{2,t} &=& \sqrt{1+t^2}, \\
 \rho_t &=& \frac{1}{\sqrt{2}}.
\end{eqnarray*}

To simulate the stochastic process $(X^1_t,X^2_t)$, we first build the model 
by setModel as follows. 
It should be noted that the method of generating 
nonsynchronous data can be replaced by a simpler one but 
we will take a general approach here to demonstrate 
a usage of 
the yuima comprehensive package for simulation and estimation of 
stochastic processes.

<<print=FALSE,echo=TRUE>>=
# diffusion coefficient for process 1
diff.coef.1 <- function(t,x1=0, x2=0) sqrt(1+t)
# diffusion coefficient for process 2
diff.coef.2 <- function(t,x1=0, x2=0) sqrt(1+t^2)
# correlation
cor.rho <- function(t,x1=0, x2=0) sqrt(1/2)
# coefficient matrix for diffusion term
diff.coef.matrix <- matrix( c( "diff.coef.1(t,x1,x2)", "diff.coef.2(t,x1,x2) * cor.rho(t,x1,x2)", "", "diff.coef.2(t,x1,x2) * sqrt(1-cor.rho(t,x1,x2)^2)"),2,2)
# Model sde using yuima.model
cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix, solve.variable=c("x1","x2"))
@

The parameter we want to estimate is the quadratic covariation between $X_1$ and $X_2$:

\begin{equation}
\theta = \langle X_1, X_2 \rangle_T =
 \int_0^T \sigma_{1,t} \sigma_{2,t} \rho_t \mbox{d} t.
\end{equation}
Later, we will compare estimated values with 
the true value of  $\theta$ given by 
<<echo=TRUE>>=
CC.theta <- function( T, sigma1, sigma2, rho)
{
 	tmp <- function(t) return( sigma1(t) * sigma2(t) * rho(t) )
 	integrate(tmp,0,T)
}
# Cumulative Covariance
theta <- CC.theta(T=1, sigma1=diff.coef.1, sigma2=diff.coef.2, rho=cor.rho)$value
cat(sprintf("theta=%5.3f\n",theta))
@ 
For the sampling scheme, we will consider the independent oisson sampling. 
That is, each configuration of the sampling times $T^{l,i}$ is realized 
as the Poisson random measure with intensity $np_l$, 
and the two random measures are independent each other as well as 
the stochastic processes. 
Then it is known that

\begin{equation}
n^{1/2} ( U_n -\theta) \rightarrow N(0,c),
\end{equation}
as $n\to\infty$, 
where 
\begin{equation}
 c = \left( \frac{2}{p_1} + \frac{2}{p_2} \right)
\int_0^T \left( \sigma_{1,t} \sigma_{2,t}  \right)^2 \mbox{d}t +
\left(
\frac{2}{p_1} + \frac{2}{p_2} - \frac{2}{p_1 + p_2}
\right)
\int_0^T
\left(
\sigma_{1,t} \sigma_{2,t} \rho_t
\right)^2 \mbox{d} t.
\end{equation}

We use a function \code{poisson.random.sampling} to get observations 
by the Poisson sampling.
This function returns a matrix combined times and observations of
Poisson random sampling.

<<echo=TRUE>>=
set.seed(111)

Terminal <- 1
p1 <- 0.2
p2 <- 0.3
rate<-c(p1, p2)
n <- 1000 
# Change to n<-1000
division <- 1200
# Change too division <- 1200

yuima.samp <- setSampling(Terminal=Terminal,division=division)
yuima <- setYuima(model=cor.mod, sampling=yuima.samp)

# solve SDEs using Euler-Maruyama method
Data <- simulate(yuima)

sampledData <- poisson.random.sampling(Data,rate=rate, n=n)
@ 
\code{cce} takes the sample and returns 
an estimate of the quadratic covariation.
<<cce, echo=TRUE>>=
cce(sampledData)[1,2] # CumulativeCovarianceEstimator
@ 
and in case of non missing data
<<cce2, echo=TRUE>>=
cce(Data)[1,2] # CumulativeCovarianceEstimator
@ 
The last value is just for reference; it may bear another kind error 
due to approximation error to the Poisson sampling.

In what follows, we will evaluate the difference between 
the estimated value and the true value, as well as the asymptotic variance, 
and compare the histogram of the estimation error after scaling 
with the limiting Gaussian distribution 
(Figure \ref{fig:histogram}). 
The function \texttt{var.c} calculates the theoretical 
asymptotic variance of the scaled estimation error.

<<cctheta2, echo=TRUE>>=
#error <- NULL   # difference between true parameter and estimator

# Cumulative Covariance
theta <- CC.theta(T=Terminal, diff.coef.1, diff.coef.2, cor.rho)$value
cat(sprintf("The true value of theta=%5.3f\n",theta))

# asymptotic variance
var.c <- function(T, p1,p2, sigma1, sigma2, rho)
{
  tmp_integrand1 <- function(t) (sigma1(t) * sigma2(t))^2
  i1 <- integrate(tmp_integrand1,0,T)
  tmp_integrand2 <- function(t) (sigma1(t) * sigma2(t) * rho(t))^2
  i2 <- integrate(tmp_integrand2,0,T)
  2*(1/p1 + 1/p2)* i1$value + 2*(1/p1+1/p2 - 1/(p1+p2)) * i2$value
}

vc <- var.c(T=Terminal, p1, p2, diff.coef.1, diff.coef.2, cor.rho)
sd <- sqrt(vc/n)
cat(sprintf("The standard deviation by normal approximation =%5.3f\n",sd))

# simulation
N <- 100
U <- NULL
for(i in 1:N)
{
  # solve SDEs using Euler-Maruyama method
  yuima.tmp <- simulate(yuima)
  yuimaData <- poisson.random.sampling(yuima.tmp,rate=rate,n=n)
  Un <- cce(yuimaData)[1,2] # CumulativeCovarianceEstimator
  U <- append(U, Un)
}

cat(sprintf("The sample mean =%5.3f\n", mean(U)))
cat(sprintf("The sample standard deviation =%5.3f\n", sqrt(var(U))))
@

\begin{figure}[h]
\centering
<<fig1, fig=TRUE,echo=TRUE,width=10,height=10,results=hide>>=
# plot sampled data histogram and Gaussian distribution (theoretical result).
library(KernSmooth)
h <- dpih(U)
bins <- seq(min(U)-5, max(U)+5+h,by=h)
xrange <- c(min(U)-5-h, max(U)+5+h)
yrange<-c(0,max(hist(U)$density))
hist(U,breaks=bins,xlab="",ylab="",xlim=xrange,ylim=yrange,probability=T)
curve(dnorm(x,mean=theta,sd=sd),ylab="",xlim=xrange,ylim=yrange,col="red", add=TRUE)
@  
\caption{A sampled data histogram and Gaussian distribution(theoretical result)}
\label{fig:histogram}
\end{figure}

\begin{thebibliography}{99}
\bibitem{hay-yos05}
Hayashi, T., Yoshida, N.: 
On Covariance Estimation of Non-synchronously Observed Diffusion Processes. 
Bernoulli, {\bf 11}, 2, 359-379 (2005)

\bibitem{hay-yos04}
Hayashi, T., Yoshida, N.: 
Asymptotic normality of a covariance estimator for nonsynchronously observed diffusion processes. 
Annals of the Institute of Statistical Mathematics, 
{\bf 60}, 2, %/ 2008”N6ŒŽ 
367-406 (2008)

\bibitem{hay-yos06}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem. 
Research Memorandum No.1020, Institute of Statistical Mathematics, 2006

\bibitem{hay-yos08}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem II. 
Research Memorandum No.1067, Institute of Statistical Mathematics, 2008

\end{thebibliography}