### 2009-11-10 pkg/yuimadocs/inst/doc/bib.Rnw
first version of yuima doc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@14 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\begin{thebibliography}{99}
\bibitem{hay-yos05}
Hayashi, T., Yoshida, N.: 
On Covariance Estimation of Non-synchronously Observed Diffusion Processes. 
Bernoulli, {\bf 11}, 2, 359-379 (2005)

\bibitem{hay-yos04}
Hayashi, T., Yoshida, N.: 
Asymptotic normality of a covariance estimator for nonsynchronously observed diffusion processes. 
Annals of the Institute of Statistical Mathematics, 
{\bf 60}, 2, %/ 2008Ç¯6·î 
367-406 (2008)

\bibitem{hay-yos06}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem. 
Research Memorandum No.1020, Institute of Statistical Mathematics, 2006

\bibitem{hay-yos08}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem II. 
Research Memorandum No.1067, Institute of Statistical Mathematics, 2008

\bibitem[Yoshida, 2005a]{Yos05a}
Yoshida, N. (2005). 
\newblock 
General M-estimation for stochastic differential equation with 
jumps by sampled data. 
\newblock 
in preparation

\end{thebibliography}