### 2009-11-11 pkg/yuimadocs/inst/doc/yuima_5.Rnw
section 5
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@15 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{adaBayes}
  adaBayes is a function that computes Bayesian estimators for unknown parameters of 
  a stochastic differential equation based on discretely observed data.

\subsection{Bayesian type estimators for the stochastic differential equation}\label{190105-4}

Consider a diffusion process $X=(X_t)_{t\in\bbR_+}$ satisfying the 
stochastic differential equation 
\beas 
dX_t=a(X_t,\theta_2)dt+b(X_t,\theta_1)dw_t,
\sskip X_0=x_0,
\eeas
where $w_t$ is an{$r$}-dimensional standard Wiener process 
independent of the initial value $x_0$. 
We suppose that the parameters 
$\theta_1$ and $\theta_2$ are unknown 
but 
$\theta_i\in\Theta_i\subset\bbR^{m_i}$ 
for $i=1,2$. 
Also assume that 
$B(x,\theta_1)=bb'(x,\theta_1)$ is elliptic uniformly 
in $(x,\theta_1)$.

In order to estimate the unknown parameters 
with the discrete-time 
observations ${\bf x}_n=(X_{t_i})_{i=0}^n$, 
$t_i=ih$ with $h=h_n$ depending on $n\in\bbN$,  
we use the quasi-likelihood function 
\beas 
p_n({\bf x}_n,\theta)
&=&\prod_{i=1}^n\frac{1}{(2\pi h)^{d/2}|B(X_{t_{i-1}},\theta_1)|^{1/2}}
\\&&
\cdot 
\exp\left(-\frac{1}{2h}
B(X_{t_{i-1}},\theta_1)^{-1}\left[
(\Delta_iX-ha(X_{t_{i-1}},\theta_2) )^{\otimes2}\right]\right),
\eeas
where $\Delta_iX=X_{t_i}-X_{t_{i-1}}$. 
%

Two approaches are possible. 
One is the quasi-maximum likelihood approach  
and another is the adaptive Bayesian approach.

The quasi-maximum likelihood estimator 
that maximizes $p_n({\bf x}_n,\theta)$ 
in $\theta=(\theta_1,\theta_2)\in
=\overline{\Theta_1\times\Theta_2}$ 
is denoted by $\hat{\theta}_{n}=
(\hat{\theta}_{1,n},\hat{\theta}_{2,n})$.

The adaptive Bayes type estimator is defined as follows. 
First we choose arbitrary value $\theta_2^\star\in\Theta_2$ and 
pretend $\theta_1$ is the unknown parameter to 
make the Bayesian type estimator $\tilde{\theta}_1$ as 
\beas 
\tilde{\theta}_1
&=&
\Big[\int_{\Theta_1}p_n({\bf x}_n,(\theta_1,\theta_2^\star))
\pi_1(\theta_1)d\theta_1 \Big]^{-1}
\int_{\Theta_1} \theta_1 p_n({\bf x}_n,(\theta_1,\theta_2^\star))
\pi_1(\theta_1)d\theta_1, 
\eeas
where 
$\pi_1$ is a prior density on $\Theta_1$. 
According to the asymptotic theory, 
if $\pi_1$ is positive on $\Theta_1$, any function can be used. 
For estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
to reform the quasi-likelihood function. That is,
the Bayes type estimator for $\theta_2$ is defined by 
\beas 
\tilde{\theta}_2
&=&
\Big[\int_{\Theta_2}p_n({\bf x}_n,(\tilde{\theta}_1,\theta_2))
\pi_2(\theta_2)d\theta_2 \Big]^{-1}
\int_{\Theta_2} \theta_2 p_n({\bf x}_n,(\tilde{\theta}_1,\theta_2))
\pi_2(\theta_2)d\theta_2, 
\eeas
where 
$\pi_2$ is a prior density on $\Theta_2$. 
In this way, we obtain the adaptive Bayes type estimator 
$\tilde{\theta}=(\tilde{\theta}_1,\tilde{\theta}_2)$ 
for $\theta=(\theta_1,\theta_2)$.

The asymptotic behavior of those estimators is known 
in various situations. The results depend on what kind 
asymptotics one considers.

See \cite{yos05}.

\subsection{Example}